\documentclass[12pt]{article}
\usepackage{amssymb}
\usepackage{amscd}
\usepackage{amsxtra}
\usepackage{enumitem}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\ep}{\varepsilon}
\newcommand{\set}[1]{\left\{ #1\right\}}
\newenvironment{proof}{\noindent{\bf Proof.}}{\hfill $\square$\medskip}

\usepackage[utf8]{inputenc}



\title{Math 108C Homework 6}
\author{Rad Mallari}

\begin{document}
\maketitle

\section{Problem 1}
Let $S$ be a symmetric matrix and let $\lambda_{1}$ and $\lambda_{2}$ be two distinct eigenvalues of $S$. Show that the corresponding eigenspaces are orthogonal.

\begin{proof}
Taking two arbitrary eigenvectors $u$ and $v$, corresponding to each eigenvalue, we have the following:
$$Su=\lambda_{1}u\quad Sv=\lambda_{2}v$$
For the eigenspaces to be orthogonal, it must be the case that $u\cdot v=0$ and
so:
\begin{equation}
    \begin{split}
        \lambda_{1}(u\cdot v)&=(\lambda_{1}u)\cdot v\\
        &=Su\cdot v=(Su)^{T}v\\
        &=u^{T}S^{T}v\\
        &=u^{T}Sv\quad\text{(by symmetry of $S$)}\\
        &=u^{T}\lambda_{2}v\\
        &=\lambda_{2}(u\cdot v) 
    \end{split}
\end{equation}
From here, we see that $\lambda_{1}(u\cdot v)=\lambda_{2}(u\cdot v)$ which implies $(\lambda_{1}-\lambda_{2})(u\cdot v)=0$. We know that $\lambda_{1}\neq\lambda_{2}\Rightarrow\lambda_{1}-\lambda_{2}=0$, therefore $u\cdot v$ must be $0$, implying they are orthogonal.
\end{proof}

\newpage
\section{Problem 5}
Show that if $S$ is positive definite, then $S$ is invertible and its inverse is also positive definite.

\begin{proof}
Since $S$ is a positive definite, we let $S=QDQ^{T}$ where $D$ are the diagonal entries. $QDQ^{T}$ here are invertible which implies that $S$ is invertible as it is a product of invertible matrices. Now 
$$S^{-1}=QD^{-1}Q^{T}$$
Is a postive definite matrix as well since $D^{-1}$ has positive diagonal entries.
\end{proof}


\section{Problem 7}
For what numbers $b$ is the following matrix positive semi-definite?
$$A=\begin{bmatrix}
    2&-1&b\\
    -1&2&-1\\
    b&-1&2
\end{bmatrix}$$

\begin{proof}
    Letting $b=2$, we have a semi-positive definite matrix.
\end{proof}
\end{document}