\documentclass[11pt]{article}
\usepackage{amssymb}
\usepackage{amscd}
\usepackage{amsxtra}
\usepackage{amsmath}
\usepackage{enumitem}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\ep}{\varepsilon}
\newcommand{\set}[1]{\left\{ #1\right\}}
\newenvironment{proof}{\noindent{\bf Proof.}}{\hfill $\square$\medskip}

\usepackage[utf8]{inputenc}


\title{Math 122B Homework 5}
\author{Rad Mallari}

\begin{document}
\maketitle

\section{Problem 1}
Find a conformal mapping of the upper half-plane onto a equilateral triangle.

\begin{proof}
Similar to \textbf{Section 13.3} in the book, we can use
\begin{equation}
    \begin{split}
        f(z)=\int_{0}^{z}f'(\zeta)d\zeta&=\int_{0}^{z}\frac{1}{(\zeta-1)^{\frac{2}{3}}(\zeta+1)^{\frac{2}{3}}}d\zeta\\
        &=\int_{0}^{z}\frac{1}{(\zeta^{2}-1)^{\frac{2}{3}}}
    \end{split}
\end{equation}
$$$$

\end{proof}


\section{Problem 2}
Suppose $f$ is an entire function mapping a rectangle to a rectangle.
Prove that $f(z)=az+b$, where $a\neq0$ and $b$ are complex numbers.


\begin{proof}

\end{proof}


\section{Problem 3}
Let $R$ be an open, simply connected subset of the complex plane. Let $z_{1},z_{2}\in R$.
Prove that there exists a conformal mapping of $R$ onto $R$ which takes $z_{1}$ into $z_{2}$.

\begin{proof}

\end{proof}


\section{Problem 4}
Let $R$ be an open, simply connected domain, different than the entire complex $\C$.
Prove that there exists no conformal mapping of $\C$ onto $R$.

\begin{proof}
Since $R$ is an open, simply connected domain, it must be the case that there exists a comformal
map $f$ from $R$ to our domain $D$ by the Riemann Mapping theorem. Say there exists a map
$g$ from $\C$ to $R$, then $f\circ g$ is a comformal map from $C$ to $D$, and since it is bounded,
it must be constant. This tells us that $g$ is not injective, thereby contradiction.
\end{proof}
\end{document}